---
title: "DA5020.A2.Katie.McCreedy"
author: "Katie McCreedy"
date: "2023-01-21"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=TRUE}
#Questions Set-up
install.packages("dplyr")
install.packages("ggplot2")
library(dplyr)
library(ggplot2)
attach(msleep)
df<-msleep
summary(msleep)
```
Preparing the dataframe, loading requisite packages, summarizing the data.


```{r}
#Question 2
table1<-table(df$vore)
library(sur)
percenttable<-percent.table(df$vore)

barplot(table1, main="Animal Eating Preferences",
   xlab="-vore status", ylab="proportion (%)")
```
The above barplot depicts the proportion of animals in the dataset by eating habit category (vore) including carnivore, herbivore, insectivore, and omnivore.

```{r}
#Question 3
omnisleep<-filter(msleep, vore =="omni")
summarise(omnisleep, avg=mean(sleep_total))
```
Filted for omnivores in the dataset and calculated the mean = 10.95. This means that for the omnivores in the dataset, the average animal gets 10.95 hours of sleep. 

```{r}
#Question 4
attach(msleep)
plot2<-plot(log10(msleep$brainwt),log10(msleep$bodywt),col = "purple", main = "brainwt vs. bodywt")
```
This scatterplot represents the linear relationship between brain weight and body weight amongst the animals in this msleep dataset. There is a strong, positively linear relationship between these variables especially after adjusting to a logarthymic scale.

```{r}
#Question 5
library(ggplot2)
plot3 <- ggplot(df, aes(log10(bodywt), log10(brainwt), colour = "blue"))+ geom_point() + geom_smooth(alpha=0.3, method="lm")
print(plot3)

library(stats)
cor(msleep$bodywt,msleep$brainwt)

#Question 6 - finding outliers
x<-msleep$sleep_total
qqnorm(x, main = "Normal Q-Q plot")
mean = mean(sleep_total)
sd = sd(sleep_total)

# threshold for outliers
xmin = mean-(1.5*sd)
xmax = mean+(1.5*sd)

# finding outliers
outliers<-(which(x < xmin | x > xmax))
outliers3<-msleep[c(10, 18, 20, 22, 23, 24, 30, 31, 36, 37, 43, 59, 62), ] 
outliers4<-outliers3[ , c("name", "sleep_total")]  
print(outliers4)
```
Practice printing the correlation plot using ggplot! Next, the pearson correlation produces "NA" value. This is because the pearson correlation test assumes that there is a linear relationship between both variables. Clearly, this is not the case, which differs from the visually seeming linear scatterplot in Q4. 

Question #6: The Normal QQ-plot first visually displays outliers in the upper right-hand corner. Then, created the thresholds for 1.5 times the standard deviation for sleep_total outside of which the outliers lay. Finally, a table is displayed presenting the outliers by their name and total sleep

