---
title: "DA5020_A6_KatieMcCreedy"
author: "Katie McCreedy"
date: "2023-02-28"
output:
  pdf_document: default
  html_document: default
---

## Question 1 - sql code
#Last login: Fri Feb 24 19:17:34 on ttys000
#katiemccreedy@Katies-MacBook-Air ~ % 
#katiemccreedy@Katies-MacBook-Air ~ % sqlite3
#SQLite version 3.37.0 2021-12-09 01:34:53
#Enter ".help" for usage hints.
#Connected to a transient in-memory database.
#Use ".open FILENAME" to reopen on a persistent database.
#sqlite> .open imdb.db
#sqlite> .read imdb.db
#Usage: .read FILE
#sqlite> .import directors.csv director_info
#sqlite> CREATE TABLE director_info (
   ...>     Director_ID INTEGER PRIMARY KEY,
   ...>     Director_Name TEXT
   ...> );
#sqlite> .mode csv
#sqlite> .import directors.csv director_info
#sqlite> SELECT*FROM director_info;
#1,"Aamir Khan"
#2,"Aaron Sorkin"


```{r setup, include=FALSE}
# Load the DBI and RSQLite packages
library(DBI)
library(RSQLite)
```

```{r echo=TRUE}
#connect to the database
con <- dbConnect(RSQLite::SQLite(), "imdb.db(2)")

# Count the number of rows in the "director_info" table
n_directors <- dbGetQuery(con, "SELECT COUNT(*) FROM director_info")$`COUNT(*)`

con <- dbConnect(RSQLite::SQLite(), "imdb.db")
# Count the number of rows in the "movie_info" table
n_movies <- dbGetQuery(con, "SELECT COUNT(*) FROM movie_info")$`COUNT(*)`
dbDisconnect(con)
```
There are 1,097 rows listed in the director_info table for 548 directors. There are 1,000 rows for movie_info. 

```{r echo=TRUE, tidy=TRUE}
# Question 2 - parts 1 & 2
library(RSQLite)
con <- dbConnect(SQLite(), "imdb.db")
num_movies <- dbGetQuery(con, "SELECT COUNT(*) as num_movies FROM movie_info")


library(RSQLite)
con <- dbConnect(SQLite(), "imdb.db")
num_movies <- dbGetQuery(con, "SELECT COUNT(*) as num_movies FROM movie_info WHERE Release_Year >= 2010 AND Release_Year <= 2020;")
#248 movies are listed between 2010-2020

# visualize the number of movies produced grouped by year
library(RSQLite)
con <- dbConnect(SQLite(), "imdb.db")
movies_by_year <- dbGetQuery(con, "SELECT Release_Year, COUNT(*) as num_movies FROM movie_info WHERE Release_Year >= 2010 AND Release_Year <= 2020 GROUP BY Release_Year;")
print(movies_by_year)

# create a bar chart with number of movies released grouped by year
library(ggplot2)
ggplot(movies_by_year, aes(x = Release_Year, y = num_movies, fill = factor(Release_Year))) +
  geom_bar(stat = "identity", color = "black") +
  ggtitle("Number of Movies Released by Year") +
  xlab("Year") +
  ylab("Number of Movies") +
  scale_fill_manual(values = c("#b39ddb", "#9575cd", "#7e57c2", "#673ab7", "#5e35b1", "#512da8", "#4527a0", "#311b92", "#1a237e", "#0d47a1", "#01579b")) +
theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
dbDisconnect(con)
```

This bar graph visualy depicts the number of movies released each year, grouped by year, between 2010-2020 with each year being depicted by a different shade of purple gradiently (I learned how to do gradient shades in one of the datacamp courses, so I wanted to practice). There were 248 total movies produced in this time period.

```{r echo=TRUE}
#Question 2 - part 3
library(RSQLite)
con <- dbConnect(SQLite(), "imdb.db")
action_ratings <- dbGetQuery(con, "SELECT MIN(IMDB_rating) AS min_rating, AVG(IMDB_rating) AS avg_rating, MAX(IMDB_rating) AS max_rating FROM movie_info WHERE Genre LIKE '%Action%';")
print(action_ratings)
dbDisconnect(con)
```
The minimum rating was 7.6, the average was 7.94, the maximum was 9 across all movies in the dataset with the genre "action". I found action movies by using wildcard operators around the word action since each genre had various specifications.

```{r echo=TRUE}
# Question 2 - Part 4
library(RSQLite)
con <- dbConnect(SQLite(), "imdb.db")
top_25 <- dbGetQuery(con, "SELECT Series_Title, Genre, CAST(Gross AS REAL) AS gross_numeric FROM movie_info ORDER BY gross_numeric DESC LIMIT 25;")
print(top_25)
dbDisconnect(con)
```
This provides the top 25 grossing movies in the dataset, ranked in descending order. Notably, Star Wars, Avengers,and Avatar, which are all series movies, are in the top 3. I had to convert gross into a numeric value so that it could order the movies.

```{r echo=TRUE}
# Question 2 - Part 5
library(RSQLite)
library(ggplot2)
library(dplyr)

# Connect to the imdb.db database
con <- dbConnect(RSQLite::SQLite(), "imdb.db")

# Read the movie_info table into a data frame
movie_info <- dbGetQuery(con, "SELECT * FROM movie_info")

# Disconnect from the database
dbDisconnect(con)

# Read the directors.csv file into a data frame
directors <- read.csv("directors.csv")

# Join the tables on Director_ID and calculate the total gross per director
result <- movie_info %>%
          inner_join(directors, by = "Director_ID") %>%
          group_by(Director) %>%
          summarize(Total_Gross = sum(Gross)) %>%
          arrange(desc(Total_Gross)) %>%
          head(10)
result

# Create a bar chart of the results
ggplot(result, aes(x = Director, y = Total_Gross)) +
  geom_bar(stat = "identity") +
  labs(title = "Directors with highest-grossing movies",
       x = "Director",
       y = "Total Gross")
```
This chart depicts the total gross of the movies produced by the top grossing movie producers including Spielberg, Jackson, and Yates. I was able to create the graph by inner-joining the Director table and the movie_info table using their common column of director_id. 

```{r echo=TRUE}
# Question 2 - Part 6
library(RSQLite)

#Write the verifyDirector function
verifyDirector <- function(Director) {
  # Read the CSV file
  directors <- read.csv("directors.csv")
  
  # Check if director was found
  if (Director %in% directors$Director) {
    message(paste0("Director '", Director, "' was found."))
  } else {
    message(paste0("Director '", Director, "' wasn't found."))
  }
  
dbDisconnect(con)
}

# Use verifyDirector to see if directors exist in the dataset
verifyDirector("Frank Darabont")
verifyDirector("Katie McCreedy")

```

The function verifyDirector checks the directors table to see whether a given director is listed in the table. Frank Darabont is returned as being in the table, because he is a movie director in the database. My name returns as "Director 'Katie McCreedy' wasn't found." because I am not a movie director. Thus, the function works. 





